---
title: "Getting Started With mlmr"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting Started With mlmr}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

## Overview

`mlmr` is an open-source Shiny app and R toolkit for fitting, understanding,
and reporting mixed-effects and multilevel models in R. It is designed for
researchers who want a guided modeling workflow while keeping the underlying R
code visible and reproducible.

The app is built around four practical goals:

- specify a model without hiding the formula;
- make centering and random-effects choices explicit;
- show equations, variance components, and Tau matrix structure;
- export tables, code, LaTeX, and Quarto-ready reports.

## Install the Public Beta

```{r install}
install.packages("remotes")
remotes::install_github("MarcusHarrisUConn/mlmr")
```

Then load the package:

```{r load}
library(mlmr)
```

## Launch the App

```{r launch}
mlmr::run_mlmr()
```

The app opens with a built-in HSB-style example so users can test the full
workflow before uploading their own data.

Uploaded data can be provided as CSV, TSV/TXT, Excel, SPSS, SAS, or Stata files.
Excel support uses the optional `readxl` package, and SPSS/SAS/Stata support
uses the optional `haven` package.

## How the Interface Is Organized

The app uses a small number of main tabs so users can move from data to model
results without losing the thread of the analysis.

- **Data** is for selecting the built-in example, uploading files, checking
  missingness, and reviewing variable roles.
- **Model** is for declaring the outcome, grouping factors, predictor levels,
  centering choices, interactions, and random-effects structure.
- **Estimate** is for fitting or refitting the model and reviewing model
  readiness before estimation.
- **Results** is for model summaries, APA-style tables, equations, diagnostics,
  and comparison tools.
- **Report & Code** is for reproducible R code, raw LaTeX, and report exports.

The public documentation site includes screenshots of these screens:

<https://marcusharrisphd.com/mlmr/>

The current supported-models scope is documented in:

```{r supported-models}
mlmr::mlm_supported_models()
```

For a longer discussion, see:

```{r supported-vignette}
vignette("supported-models", package = "mlmr")
```

## Recommended First Workflow

1. Open the app.
2. Keep the example dataset selected on the **Data** tab.
3. Review grouping factors and variable roles.
4. Move to **Model** and inspect the preset outcome, predictors, centering, and
   random-effects choices.
5. Click **Fit Example Model** from **Data > Overview**, or move to
   **Estimate** and click **Fit Example Model**.
6. Open **Results** to review the dashboard, APA tables, equations,
   interpretation, diagnostics, and model comparison panels.
7. Open **Report & Code** to export reproducible R code, raw LaTeX, APA tables,
   or a Quarto-ready report.

## What the Example Demonstrates

The built-in example includes a student outcome, student-level predictors,
school-level predictors, categorical predictors, interactions, and random
effects. It is intentionally more complex than a minimal random-intercept model
so users can see why mixed-effects and multilevel models are useful.

The example currently demonstrates:

- cluster-mean centering for lower-level predictors;
- grand-mean centering for contextual predictors;
- dummy coding for categorical predictors;
- random intercepts and random slopes;
- fixed interactions and cross-level interaction-style terms;
- APA-style tables;
- level-by-level equations and combined equations;
- Tau variance-covariance matrix display.

## Reproducibility

Every model fit in `mlmr` is paired with reproducible code. The **Report & Code**
tab exports the analysis code and the raw LaTeX used for tables and equations.

The long-term goal is that users can move from the graphical workflow to a
scripted R workflow without losing any modeling decisions.

## PDF Manual and Help Pages

Like other R packages, `mlmr` has function help pages that can be viewed with:

```{r help}
?run_mlmr
help(package = "mlmr")
```

CRAN automatically creates a PDF reference manual from those help pages after a
package is accepted. During beta development, the pkgdown website and vignettes
are the main user-facing manuals because they explain the workflow in context.

## Current Status

`mlmr` is in public beta. The current version is suitable for structured demo
testing, teaching, and feedback. Users should still independently verify model
specification, convergence, diagnostics, and interpretation before using fitted
results in production research.
