---
title: "Supported Models and Production Scope"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Supported Models and Production Scope}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

## Purpose

`mlmr` is designed to make mixed-effects and multilevel models easier to
specify, understand, teach, and report. The app is intentionally opinionated:
it guides users through a declared modeling workflow while still exposing the
underlying `lme4` formula, reproducible R code, equations, diagnostics, and
manuscript-ready outputs.

This article states the current software scope plainly so users know what is
ready for routine use, what is advanced or experimental, and what remains future
work.

## Current Production Path

The strongest current workflow is a Gaussian mixed-effects model for nested
educational, psychological, health, or organizational data. The app is
especially well suited for:

- two-level nested models, such as students within schools;
- three-level nested models, such as students within schools within districts;
- random-intercept and random-slope models;
- correlated or independent random-effects structures;
- grand-mean and cluster-mean centering;
- fixed interactions and cross-level interaction-style models;
- APA-style tables, equations, Tau matrices, diagnostics, and reproducible
  exports.

The built-in HSB-style example is intentionally more complex than a minimal
random-intercept model so users can see how centering, interactions, random
slopes, and variance-covariance structures work together.

## Supported Scope Table

The same scope table is available programmatically:

```{r scope}
mlmr::mlm_supported_models()
```

This table is useful for instructors, analysts, and beta testers because it
separates software support from statistical responsibility. A model can be
available in the interface and still require careful evaluation of assumptions,
sample size, convergence, and interpretation.

## Advanced Workflows

`mlmr` includes advanced pathways for GLMMs and flexible random-effect
structures because real research often moves beyond strict nesting. These
workflows are valuable, but they require more careful checking:

- binomial, Poisson, negative binomial, and Gamma GLMMs should be checked for
  outcome coding, link function, overdispersion, and convergence;
- crossed random effects should be used only when the grouping variables truly
  represent non-nested sources of variation;
- longitudinal models can be represented through grouping variables and time
  predictors, but users should evaluate residual patterns and time structures;
- weighted multiple-membership models are not a current production feature.

## What mlmr Does Not Replace

`mlmr` does not replace statistical judgment. It helps users express and report
models more clearly, but users remain responsible for:

- deciding whether a mixed-effects model is appropriate for the design;
- checking whether grouping factors and predictor levels are correctly declared;
- choosing centering and contrast coding that match the research question;
- evaluating convergence, singular fits, residual patterns, and influential
  observations;
- confirming that manuscript language matches the fitted model.

## Production Readiness Standard

For a model to be treated as production analysis, users should be able to:

1. reproduce the fitted model from the exported R code;
2. explain the centering, interactions, and random-effects choices;
3. inspect the fixed effects, variance components, ICCs, and Tau matrix;
4. review diagnostics and convergence messages;
5. export manuscript tables, equations, and software citations;
6. save a Quarto or script-based record of the full workflow.

That standard is the guiding target for `mlmr` development.
